---
title: "🗺️ Interactive Mapping with mapgl and Shiny"
author: "DS421 – Cartographic Design"
format:
  html:
    code-fold: true
    toc: true
    theme: cosmo
execute:
  echo: true
  warning: false
  message: false
---

## 1️⃣ Census Shiny App

#### Load Libraries

```{r}
library(shiny)
library(tidyverse)
library(sf)
library(mapgl)
library(lubridate)
library(bslib)
library(here)
library(tidycensus)
library(ggpubr)
library(plotly)
```

#### View Census Variables

```{r}
acs_vars <- load_variables(year = 2023, dataset = "acs5", cache = TRUE)
```

#### Transform Data

```{r}
# Define raw ACS variable codes
raw_vars <- c(
  median_income = "B19013_001",
  bachelors_count = "B15003_022",  # Count, not percent
  median_gross_rent = "B25064_001",
  total_population = "B01003_001"
)

# Labels for Shiny dropdown (key = column name in wide data)
oahu_choices <- c(
  "Median Household Income" = "median_income",
  "Percent with Bachelor's Degree" = "pct_bachelors",
  "Median Gross Rent" = "median_gross_rent"
)

# Pull and transform ACS data
hawaii_data_raw <- get_acs(
  geography = "tract",
  state = "HI",
  variables = raw_vars,
  geometry = TRUE,
  year = 2023
) |>
  filter(GEOID != "15003981200") |>  # Remove Kalawao County
  st_transform(4326)

# Convert to wide format + calculate % with bachelor's
hawaii_data <- hawaii_data_raw |>
  select(GEOID, NAME, variable, estimate, geometry) |>
  st_drop_geometry() |>   # temp drop for pivot
  pivot_wider(names_from = variable, values_from = estimate) |>
  mutate(
    pct_bachelors = 100 * bachelors_count / total_population
  ) |>
  left_join(
    hawaii_data_raw |> select(GEOID, geometry) |> distinct(),
    by = "GEOID"
  ) |>
  st_as_sf()
```

#### Build Shiny App

```{r}
ui <- page_sidebar(
  title = "Hawaii Census Map",
  sidebar = sidebar(
    selectInput("var", "Choose a variable to map:", choices = oahu_choices)
  ),
  card(
    full_screen = TRUE,
    maplibreOutput("map")
  )
)

server <- function(input, output, session) {
  output$map <- renderMaplibre({
    # 1. Get the selected column from hawaii_data
    selected_data <- hawaii_data |> 
      mutate(var_value = .data[[input$var]])

    # 2. Use interpolate_palette() with a good break method
    continuous_scale <- interpolate_palette(
      data = selected_data,
      column = "var_value",
      method = "equal",  # Or "quantile"
      n = 4,              # Reduce to 4 breaks to avoid crowding
      palette = viridisLite::viridis
    )

    # 3. Format legend labels using prefix/suffix
    legend_labels <- switch(input$var,
      "median_income" = get_legend_labels(continuous_scale, digits = 0, prefix = "$"),
      "pct_bachelors" = get_legend_labels(continuous_scale, digits = 0, suffix = "%"),
      "median_gross_rent" = get_legend_labels(continuous_scale, digits = 0, prefix = "$"),
      get_legend_labels(continuous_scale, digits = 0)  # fallback
    )

    # 4. Map with scale expression and legend
    maplibre(style = carto_style("positron")) |>
      fit_bounds(selected_data) |>
      add_fill_layer(
        id = "tracts",
        source = selected_data,
        fill_color = continuous_scale$expression,
        fill_opacity = 0.75
      ) |>
      add_legend(
        legend_title = names(oahu_choices)[oahu_choices == input$var],
        values = legend_labels,
        colors = get_legend_colors(continuous_scale),
        type = "continuous"
      )
  })
}


shinyApp(ui, server)

```

#### ggplot

Let's see if there is a relationship between our census variables by creating a scatter plot

```{r}
scatter_plot <- ggplot(hawaii_data, aes(x = median_income, y = median_gross_rent)) +
  geom_point() +
  geom_smooth() +
  stat_cor(output.type = "text", label.sep = '\n') +
  theme_minimal()

scatter_plot
```

Now lets make it interactive

```{r}
ggplotly(scatter_plot)
```

## 🧩 Challenge 1 

Can you add the functionality to your shiny app (new panel) that allows users to pick two census variables to plot on an interactive scatter plot using `ggplotly()` and `stat_cor()`

```{r}

```

## 🧩 Challenge 2

Can you create a shiny app using your LANDSAT_USA_contiguous_and_Hawaii_7d.csv in the data folder that allows users to filter the map by confidence. For example, the user can select from a side dropdown to only show *high* confidence fires, *medium* confidence fires, or *low* confidence fires. Also add a date slider so users can toggle the map by date. Bonus: Can you add a heatmap layer users can toggle on and off?

```{r}

```
